Positional entropy in childes

### 2018-08-16 positional_entropy_childes.Rmd
mutual information

compute estimates
```{r calc_pos_entropy}
get_entropies <- function(role, sen_length, utterances) {

@@

  tokens <- sub_utterances %>%
    filter(length == sen_length) %>%
    mutate(utterance_id = 1:n()) %>%
    unnest_tokens(word, stem) %>%
    group_by(utterance_id) %>%
    mutate(word_order = 1:n())

  tokens %>%
    group_by(word_order, word) %>%
    summarise(n = n()) %>%
    tidyboot(summary_function = function(x) x %>% 
               summarise(entropy = entropy(n, unit = "log2")),
             statistics_functions = function(x) x %>%
             summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
    mutate(role = role, length = sen_length)
}

entropies <- map(2:10, ~get_entropies("not child", .x, prov_utterances)) %>%
  bind_rows()

>>>>

compute estimates
```{r calc_pos_entropy}
get_counts <- function(role, sen_length, utterances) {

@@

  sub_utterances %>%
    filter(length == sen_length) %>%
    mutate(utterance_id = 1:n()) %>%
    unnest_tokens(word, stem) %>%
    group_by(utterance_id) %>%
    mutate(word_order = 1:n())
}

unigram_entropy <- function(counts) {

  counts %>%
    group_by(word_order, word) %>%
    summarise(n = n()) %>%
    tidyboot(summary_function = function(x) x %>% 
               summarise(entropy = entropy(n, unit = "log2")),
             statistics_functions = function(x) x %>%
             summarise_at(vars(entropy), funs(ci_upper, ci_lower))) %>%
    mutate(role = role, length = sen_length)
}

counts <- map(2:6, ~get_counts("not child", .x, prov_utterances))

all_pairwise_mi <- function(df) {
  max_len <- max(df$word_order)

  combs <- combn(1:max_len, 2, simplify = F)

  map(combs, ~position_mi(df, .x[1], .x[2])) %>%
    bind_rows()

}

position_mi <- function(df, pos1, pos2) {

  mi <- df %>%
    filter(word_order %in% c(pos1, pos2)) %>%
    mutate(word_order = factor(word_order, labels = c("first", "second"))) %>%
    spread(word_order, word) %>%
    group_by(first, second) %>%
    summarise(n = n()) %>%
    spread(second, n, fill = 0) %>% 
    ungroup() %>%
    select(-first) %>%
    mi.empirical(., unit = "log2")

  data_frame(length = max(df$word_order), pos1 = pos1, pos2 = pos2, mi = mi)
}

mi <- map(counts, all_pairwise_mi)

ggplot(mi, aes(x = pos2, y = mi, color = as.factor(pos1))) + 
  facet_wrap(~length) + 
  geom_point() + 
  geom_line()

ggplot(mi, aes(x = pos1, y = mi, color = as.factor(pos2))) + 
  facet_wrap(~length) + 
  geom_point() + 
  geom_line()

entropies <- map(2, ~ get_counts("not child", .x, prov_utterances) %>%
                   bind_rows %>%
                   unigram_entropy) %>%
  bind_rows()
